Simple connectome inference from partial correlation statistics in calcium imaging

### 2014-06-19 article.tex
second pass on metric choice

%\textcolor{red}{SEUIL 0.11 for tabulars!!!}

% In this section, we will first describe datasets and methods used for the
% comparison. Then, on Table \ref{tab:tab3}, our approach is decomposed to show
% the contribution of each pre-processing steps. Afterwards, we will perform a comparison
% between several methods on Figure \ref{fig:curves} and Table \ref{tab:tab1}. Note that two
% metrics are used to evaluate the performance and a single low-pass filter is
% used (Eq. \ref{eq:symetric-median}) and a threshold, i.e. $\tau = 0.11$. The end
% of this section is devoted to a discussion about the choice of metric and its
% relevance for the challenge.

% \subsection{Assessment methodology}

@@

\subsection*{Comparison with state-of-the art methods}

Figure \ref{fig:curves} shows the ROC curve and precision-recall curve
for correlation, GTE and partial correlation on the \textcolor{red}{normal-?}
dataset. The partial correlation outperforms other state-of-the-art methods
by at least factor of 2 in AUPRC and an increased of $5\%$ in AUROC. There
are two reasons that explains such overwhelming difference: (i) among
those methods only the partial correlation make the distinction between
direct and indirect connection, (ii) the simulated network from the
\citep{stetter2012model} generator could be reduced to a linear model after
pre-processing.

\begin{figure}[bth]
\centering
\includegraphics[width=0.6\linewidth]{images/curves}
\caption{Method precision-recall and roc curves. \textcolor{red}{GENI3?}}
\label{fig:curves}
\end{figure}

Observing Table \ref{tab:tab1}, it appears that best methods only infer the
\textbf{undirected} network by giving the same weights to both edges $(i,j)$
and $(j,i)$. This is quite surprising considering the goal of the challenge,
namely predicted directed connections between neurons. Besides, if the true
network undirected, i.e. for all links $(i,j)$, back links are added, is the
infered network then the score obtained according to the ROC AUC metric is
very high ($\approx 0.9956$ on \textit{normal-1}). Therefore, an undirected
infered network seems sufficient to maximize the ROC AUC score without trying
to identify the link direction. Indeed, predicting perfectly directivity only
yields a small gain ($\approx 0.005$ on \textit{normal-1}) which is is not
very rewarding. In contrast, the same experiment with the precision recall
metric rewards less the undirected network ($\approx 0.789$ on
\textit{normal-1}) and the gain margin in predicting the link directivity is
large enough to be interesting ($\approx 0.211$ on \textit{normal-1}). For
these reason, it seems that the metric used in the contest, i.e. the ROC AUC,
does not seem consistent with the goal of predicting directed connections
between neurons. Another metric, such as the AUC PRC, is maybe more
appropriate to assess the quality of the \textbf{directed} infered network.

Interestingly, an easy way to improve results for asymmetric methods (see
Tab.\ref{tab:tab2}), namely GTE and GENIE3, is to make undirected the infered
network. It proves once again that trying to detect the link direction is more
penalizing than rewarding.

% TODO
%   - discussions about the results
%   - discussions about the metrics (Gilles' text)

% Dummy classifier?

% GTE : Explaning the method (default parameters!)

% Explaining the method (thèse de Gilles)

% Directivity : ?

% Opt: obviously when parameters are tuned/optimized, it's slightly better -->
% see appendix for a full description of the method
% Avg.: Averaging over the threholds and L-P filters improves a lot both metrics.

% Partial correlation : Obvious improvements (both metric) because of the conditioning
% Correlation : Already give good results

% Averaging

@@

% Discussion about the metrics

@@

@@

@@

>>>>

@@

\subsection{Partial correlation compare favorably with other state-of-the art methods}

Figure \ref{fig:curves} shows the ROC curve and precision-recall curve
for correlation, GTE and partial correlation on the \textcolor{red}{normal-?}
dataset. The partial correlation outperforms other state-of-the-art methods
by at least factor of 2 in AUPRC reaching $0.347$ and an increased of $0.05$ in
AUROC.

Two reasons might explain such overwhelming difference: (i) among
those methods only the partial correlation makes the distinction between
direct and indirect connections, (ii) the simulated network from the
\citep{stetter2012model} generator could be reduced to a linear model after
pre-processing. \textcolor{red}{Those results generalized over the other
datasets see Appendix REF.}

@@

% Opt: obviously when parameters are tuned/optimized, it's slightly better -->
% see appendix for a full description of the method

\begin{figure}[bth]
\centering
\includegraphics[width=0.6\linewidth]{images/curves}
\caption{Method precision-recall and roc curves. \textcolor{red}{GENI3?}
         $\tau=0.11$}
\label{fig:curves}
\end{figure}

@@

% Explaining the method (thèse de Gilles)
% Avg.: Averaging over the threholds and L-P filters improves a lot both metrics.
% Averaging

@@

\subsection{Area under the ROC curve is a poor metric for network inference}

The best method only infers the \textbf{undirected} network by giving the
same connectivity score to both edges $(i,j)$ and $(j,i)$. This is quite
surprising considering the problem statement.

If one would infer perfectly the undirected network, it would already yield
high AUROC score for instance $\approx{}0.996$ on \textit{normal-1}. Trying
to infer edge directivity is likely to be counter-productive or yield
small AUROC gain. The area under the ROC curve is not a consistent metric
with respect to the problem statement.
By contrast, the same experiment with AUPRC would yield
to a score of $\approx 0.789$ on \textit{normal-1}. Thus inferring
correctly directivity would be more rewarding in AUPRC term.
Interestingly, methods that takes into account directivity, namely GTE and
GENIE3, improve their AUROC and AUPRC by giving the same way to directed edges.
Trying to detect the link direction is more penalizing than rewarding.

For these reasons, we strongly advice to avoid the AUROC
for the assessment network inference using a more appropriate
metric such as the area under the precision-recall curve (AUPRC). Those
conclusions are in line with \cite{schrynemackers2013protocols}.

@@

\section{Extended results}
